\documentclass{article}
\usepackage{html}
\pagestyle{empty}

\externallabels{../../user/Utility}{../../user/Utility/labels.pl}
\externallabels{../../user/General}{../../user/General/labels.pl}
\externallabels{../../user/Display}{../../user/Display/labels.pl}
\externallabels{../../user/Dish}{../../user/Dish/labels.pl}
\externallabels{../../user/SynthesisRef}{../../user/SynthesisRef/labels.pl}
\externallabels{../../user/ATNF}{../../user/ATNF/labels.pl}
\externallabels{../../user/NFRA}{../../user/NFRA/labels.pl}
\externallabels{../../user/NRAO}{../../user/NRAO/labels.pl}
\externallabels{../../reference/Glish}{../../reference/Glish/labels.pl}

\begin{document}


{\large \textbf{Frequenty Asked Questions About Using \textsf{AIPS++}}}


\begin{enumerate}

\item {\htmlref{When I type a command, all that happens is that I see line after line of \textit{Glish} code.}{Question 1:}}

\item {\htmlref{When I type \texttt{Imager()}, I get an error.}{Question 2:}}

\item {\htmlref{When I type \texttt{x=1}, \textit{Glish} complains about error, parse error at or near \texttt{'='}.}{Question 3:}}


\item {\htmlref{When I try to set a string, \textit{Glish} complains.}{Question 4:}}
 

\item {\htmlref{I get too much output if I mistakenly type the name of an array.}{Question 5:}}

\item {\htmlref{Strange things happen if I use double quotes around strings.}{Question 6:}}

\item {\htmlref{When I start \textsf{AIPS++} and type ahead, strange things happen.}{Question 7:}}

\item {\htmlref{Some commands don't do what I expect: \texttt{doit(shift=t)}.}{Question 8:}}

\item {\htmlref{How do I run something in the background?}{Question 9:}}

\item {\htmlref{How do I stop something?}{Question 10:}}

\item {\htmlref{AIPS++ is very sloooowwwwww...}{Question 11:}}

\item {\htmlref{How do I copy or rename an Image or MeasurementSet?}{Question 12:}}

\item {\htmlref{How do I make notes in the log?}{Question 13:}}

\item {\htmlref{How can I customize AIPS++ on startup?}{Question 14:}}

\item {\htmlref{The display looks empty.}{Question 15:}}

\item {\htmlref{How do I change directory in Glish?}{Question 16:}}

\item {\htmlref{Why is the Glish prompt sometimes \texttt{-} and sometimes \texttt{+}?}{Question 17:}}

\item {\htmlref{Why do the contents of AIPS++  windows
sometimes behave strangely after the window is resized?}{Question 18:}}

\item {\htmlref{My display seems to have frozen up, but the Command Line window is still working.}{Question 19:}}

\item {\htmlref{AIPS++ just hangs, now what?.}{Question 20:}}

\end{enumerate}



\begin{description}

\item[Q:1
\label{Question 1:}When I type a command, all that happens is that I see line after line of Glish  code.]

Remember to put brackets () after the command. The name of the
function alone is a valid \textit{Glish} variable that can be passed around
just as any other variable. We use this capability extensively in the
\textit{Glish} code.

\vspace{.2in}

\item[Q:2
\label{Question 2:}When I type \texttt{Imager()}, I get an error.]

\textsf{AIPS++} is case-sensitive, so in this example you need to type
\texttt{imager()}.  The convention is that names of functions should always be
lower case, and that different words in a composite name run into each
other without, for examples, underscores to delineate them.


\item[Q:3 
\label{Question 3:}When I type \texttt{x=1}, \textit{Glish} complains about error, parse error at or near \texttt{'='}.]

Remember that \textit{Glish} needs a colon for assignment: \texttt{x:=1}. The single
equals is reserved for default arguments. (No, we don't particularly
like this feature either). 

\item[Q:4
\label{Question 4:}When I try to set a string, \textit{Glish} complains.]


Remember that \textit{Glish}  needs (single) quotes around strings. 

\begin{verbatim}
     - file:=3C84.MS                             
                  # Incorrect
     error, parse error at or near 'C84'
     warning, uninitialized global variable MS used
     F 
     - file:='3C84.MS'                           
                  # Correct

\end{verbatim}

\item[Q:5
\label{Question 5:}I get too much output if I mistakenly type the name of an array:]

\textit{Glish} assumes that you really want to see what you type. So if you
type an array name then the whole thing can be dumped to the
screen. You can limit the output per array: 

\begin{verbatim}
     a:=array(0.0, 100, 100)
     a::print.limit:=10
\end{verbatim}

     Or, more usually, for all output: 
\begin{verbatim}
     system.print.limit:=10
\end{verbatim}

\item[Q:6
\label{Question 6:}Strange things happen if I use double quotes around strings.]

The double quote tells \textit{Glish} to split the enclosed string into a
vector of strings, one for each full word, whereas single quotes
preserve the entire string as one element. This can affect tests for
equality in \textit{Glish} code. We try to avoid this error in our code but to
be safe you should always use single quotes. 

\item[Q:7
\label{Question 7:}When I start AIPS++ and type ahead, strange things happen.]

Wait until you see a - prompt. \textit{Glish} processes include files in the
background and can accept inputs while doing so although the result
may not be reasonable. So it's best to be polite and wait. Here is an
example of what can happen: 

\begin{verbatim}
     aips++
     x:=1
     "logger.g", line 330: warning$: 
      no events have been received
     "logger.g", line 330: error, 
      non-string type in coercion of <fail>
\end{verbatim}

\item[Q:8
\label{Question 8:}Some commands don't do what I expect: \texttt{doit(shift=t)}.]

The documentation says to use \texttt{shift=T}. \textbf{t} is not the same as \texttt{T} (the
true symbol in \textit{Glish}).In fact, since t is undefined, it is created and
assigned a value of \texttt{F} (the False symbol), thus doing the opposite of
what you expected! 

\item[Q:9
\label{Question 9:}How do I run something in the background?]

Some \texttt{tools} support this via the global \texttt{dowait} variable. See for
example, Image or Imager. The GUI also allows this.

\item[Q:10
\label{Question 10:}How do I stop something?]

Control-\texttt{c} will stop a \textit{Glish} function from running. GUI functions can
be stopped by pressing the \textsf{abort} button. 

\item[Q:11
\label{Question 11:}AIPS++ is very sloooowwwwww...]

\textsf{AIPS++}  is designed to run well on a canonical machine which is a Sparc
4 or Ultra or 200 MHz Pentium Pro class machine with 64 Mbytes of
memory or more. It will run very slowly on a 32MByte Sparc IPX, for
example, swapping a lot. The best way to check memory availability is
to use a system utility such as top on Solaris machines. 

\item[Q:12
\label{Question 12:}How do I copy or rename an Image or MeasurementSet?]

For any Table, you can use \texttt{tablecopy} or \texttt{tablerename}. The \texttt{Catalog}
tool will also do this, with the option of using a GUI. 


\item[Q:13
\label{Question 13:}How do I make notes in the log?]

Use the logger.note. 

\begin{verbatim}
     dl.note('Forgot robust weighting - 
     repeating deconvolution')
\end{verbatim}

\item[Q:14
\label{Question 14:}How can I customize AIPS++ on startup?]

There are two possibilities: 

\begin{verbatim}
     .glishrc 
\end{verbatim}

         A \textit{Glish} defaults file, \textit{.glishrc}, can be located in your home
         directory. The \textit{.glishrc}  file can be used to customize \textit{Glish},
         such as to set search paths for \textit{Glish} scripts, to set default
         precision on output, to start particular packages, etc. It
         can contain any valid \textit{Glish} command. The two most important
         uses are tell \textit{Glish} to use a particular directory for the
         search path for include files, and to limit the amount of
         output that \textit{Glish} sends to the screen:

\begin{verbatim}         
system.path.include:=
          '. /home/tarzan/tcornwel/aips++/libexec 
/aips++/beta/sun4sol_gnu/libexec'
system.print.limit:=10
\end{verbatim}

         \textit{.aipsrc} An \textsf{AIPS++} defaults file, \textit{.aipsrc}, can be located in
         your home directory. This is used to hold global choices and
         information.

         For example:
 
\begin{verbatim}
catalog.default: gui          
          #start a GUI catalog on startup
		
measures.default: gui       
          #start a GUI measures on startup
          
logger.default: screen       
          #use the screen for Logger messages
          
aipsview.numcolors: 200   
          #use 200 colors in Aips View palette
\end{verbatim}


\item[Q:15
\label{Question 15:}The display looks empty.]

Perhaps Netscape has gobbled up all the color table. You have two
choices, either stop Netscape and restart it with the -install
command-line option to require a private color map, or specify aipsrc
variables to start Aips View with a smaller number of colors (see
display). 


\item[Q:16
\label{Question 16:}How do I change directory in \textit{Glish}?]

You can't at the moment. Instead, you'll have to exit from \textit{Glish}, \texttt{cd}
at the system prompt, and then restart \textsf{AIPS++}. An alternative is use
file names with directories specified. We will add capabilitity to
change directory in the future. 

\item[Q:17
\label{Question 17:}Why is the Glish prompt sometimes \texttt{-} and sometimes \texttt{+}?]

The statement that you entered is ambiguous. Usually entering a simple
semi-colon will resolve it.

\item[Q:18
\label{Question 18:}Why do the contents of the AIPS++
windows sometimes behave strangely after the window is resized?]

It is a "feature" of the Tk widget set that we use.  Once any window is resized, the new size is
taken to be fixed, and any subsequent changes to the contents are constrained to fit within
the window size, instead of expanding it as necessary.  The workaround is not to resize 
windows unless necessary.

\item[Q:19
\label{Question 19:}My display seems to have frozen up, but the Command
Line window is still working.]

A logic error in a GUI can result in this.  First, at the Command Line window, type:

\begin{verbatim}
tk_release()
\end{verbatim}

Try this a few times to see if the displays unfreeze.  If so, then please submit a bug 
report if you think you know under what conditions the freeze occurred.

\item[Q:20
\label{Question 20:}AIPS++ just hangs, now what?.]
Check that lockd is running on your NFS server.
\end{description}

Please E-mail any comments or questions about \textsf{AIPS++} to
\texttt{aips2-request@nrao.edu.}

\textit{Copyright 1995-2000,2002 Associated Universities Inc.,
Washington, D.C.}
\end{document}
