%% Copyright (C) 1999,2000,2001,2002,2003
%% Associated Universities, Inc. Washington DC, USA.
%%
%% This library is free software; you can redistribute it and/or modify it
%% under the terms of the GNU Library General Public License as published by
%% the Free Software Foundation; either version 2 of the License, or (at your
%% option) any later version.
%%
%% This library is distributed in the hope that it will be useful, but WITHOUT
%% ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE.  See the GNU Library General Public
%% License for more details.
%%
%% You should have received a copy of the GNU Library General Public License
%% along with this library; if not, write to the Free Software Foundation,
%% Inc., 675 Massachusetts Ave, Cambridge, MA 02139, USA.
%%
%% Correspondence concerning AIPS++ should be addressed as follows:
%%        Internet email: aips2-request@nrao.edu.
%%        Postal address: AIPS++ Project Office
%%                        National Radio Astronomy Observatory
%%                        520 Edgemont Road
%%                        Charlottesville, VA 22903-2475 USA
%%
%% $Id$
\NextFile{sd}
\begin{ahmodule}{sd}{Single-dish data analysis package}

\begin{ahdescription}

The sd module provides functions and tools to reduce and analyse 
single-dish data. 
The module can be categorised as 

\begin{itemize}

\item \ahlink{sd}{sd:sd} - Top level tool that contains the following sub 
tools and general functions for single-dish data reduction 
and analysis such as calibration, data averaging, etc. 

\item \ahlink{sd.scantable}{sd:sd.scantable} - Representation of single-dish data 
format. The tool also contains getter/setter functions as well as functions 
that manipulates data contents such as baseline fitting, flagging, etc. 

\item \ahlink{sd.selector}{sd:sd.selector} - Data selector that operates on 
scantable.

\item \ahlink{sd.fitter}{sd:sd.fitter} - This tool is an engine to perform both 
baseline and spectral line fitting.

\item \ahlink{sd.linefinder}{sd:sd.linefinder} - This tool performs automatic 
line finding according to the user specified thresholds.

\item \ahlink{sd.linecatalog}{sd:sd.linecatalog} - Line catalog. The module has 
its own line catalog for this tool. The tool also accepts user provided 
catalog in the form of ASCII table with specific format.

\item \ahlink{sd.plotter}{sd:sd.plotter} - Plotter tool dedicated for 
single-dish data (scantable)

%\item \ahlink{interactivemask}{sd:sd.interactivemask} - Tool for GUI based 
%region selection.

\item \ahlink{sd.coordinate}{sd:sd.coordinate} - Tool for conversion from pixel to 
velocity or frequency.

%\item \ahlink{lagflagger}{sd:sd.lagflatter} - 

\item \ahlink{sd.opacity\_model}{sd:sd.opacity_model} - MIRIAD like opacity model.

\end{itemize}


\end{ahdescription}

\ahobjs{}
\ahfuncs{}

\input{sd.htex}

\end{ahmodule}
