\documentclass{article}
\usepackage{html}
\pagestyle{empty}

\externallabels{../../user/Utility}{../../user/Utility/labels.pl}
\externallabels{../../user/General}{../../user/General/labels.pl}
\externallabels{../../user/Display}{../../user/Display/labels.pl}
\externallabels{../../user/Dish}{../../user/Dish/labels.pl}
\externallabels{../../user/SynthesisRef}{../../user/SynthesisRef/labels.pl}
\externallabels{../../user/ATNF}{../../user/ATNF/labels.pl}
\externallabels{../../user/NFRA}{../../user/NFRA/labels.pl}
\externallabels{../../user/NRAO}{../../user/NRAO/labels.pl}
\externallabels{../../reference/Glish}{../../reference/Glish/labels.pl}
\begin{document}


{\large \textbf{Frequently Asked Questions About AIPS++}}

\begin{enumerate}
\item \textit{What is \textsf{AIPS++}}?

\textsf{AIPS++} is the Astronomical Information Processing System.


\item \textit{How was it developed?}

By a \htmladdnormallink{Consortium}{../../project/consortium.html} of radio
observatories. About 40 people in all contributed to the first release
in 1999. The design was based on object-oriented techniques, and the
implementation was in \textbf{C++} and a glue language called 
\htmladdnormallink{Glish}{../../glish/glish.html}


\item \textit{What does it do?}

We quote the \htmladdnormallink{mission statement}{../../project/mission.html}

\textit{A telescope requires post-processing software for calibration,
editing, image formation, image enhancement, and analysis of images
and other data streams. This software is an integral part of the radio
telescope engineering. The Astronomical Information Processing System
(\textsf{AIPS++}) project is designed to produce such a software product.}

\textit{Although \textsf{AIPS++} is primarily targeted at radio astronomy, 
it is
anticipated that it will also be used in other branches of astronomy
and for other applications in image processing and data analysis.}

\item \textit{How is it related to \htmladdnormallink{AIPS}
{http://www.cv.nrao.edu/aips/}?}

\textsf{AIPS++} is the next generation. Little is in common
with AIPS, besides some of the algorithms. Data structures,
command language, algorithms, user interfaces, applicability
have all been substantially improved relative to AIPS.

\end{enumerate}


If you find that the response is slow, documents are out of
date, or you receive an error message,  it may help to have documentation served to you from a \htmladdnormallink{mirror site}{../../html/mirrors.html}

Please E-mail any comments or questions about \textsf{AIPS++} to
\texttt{aips2-request@nrao.edu}.






Copyright 1995,1996,1997,1998,1999,2000 Associated Universities Inc.,
Washington, D.C.
  
\end{document}
